---
title: "Model Fit - Stan - V1"
author: "Ben Margetts"
date: "02/02/2018"
output:
  pdf_document: default
  html_document: default
---

RUNS Model on data from ODESimulations and ODESimulationsMyModel.Rmd

# Description

Nowak model with basic functional code.
V1 = reasonable estimates, working. Good Rhat values

V1.5 = :
  - y0s
  - t0s > 0
  - Generated quantities

# Libraries
```{r}
library(rstan)
library(ggplot2)
library(scales)
library(plyr)
#library(ctsem)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Function to kill Stan runs (with multiple chains)
```{r}
kill_stan = function(){
	system2(
		command = "killall"
		, args = "R"
	)
}
```


# Working Directory
```{r}
wd <- '~/Google/PhD/Year_1/Modelling/Stan/CMV Stan Feb 18/'
```

# Read Data In
```{r}
setwd(wd)
modelDat <- read.csv('ModellingData.csv', stringsAsFactors = F)

#TEST
#modelDat <- modelDat[modelDat$Event == 1 | modelDat$Event == 2 | modelDat$Event == 7,]
modelDat <- modelDat[modelDat$Event %in% sample(unique(modelDat$Event), size = 10),]
```

# Model parameters
```{r}
t0 <- -1
numODEs <- 3
numIDs <- length(unique(modelDat$Event))

Y0s <- vector()
for (i in unique(modelDat$Event)){
  #Y0s <- append(Y0s, modelDat$Y0[modelDat$ID==i][1])
  Y0s <- append(Y0s, unique(modelDat$Y0[modelDat$Event == i]))
}

# Adjust for t0
#modelDatUse <- modelDat[!modelDat$Time == 0,]
modelDatUse <- modelDat
#modelDatUse$Time <- modelDatUseTime + 1

numObs <- vector()
for (i in unique(modelDatUse$Event)){
  numObs <- append(numObs,length(modelDatUse$Event[modelDatUse$Event==i])) #test
}

totNumObs <- sum(numObs)
```


# Stan - Ragged Array
```{r}
sizes <- vector()

for (i in unique(modelDatUse$Event)){
  sizes <- append(sizes, length(modelDatUse$Event[modelDatUse$Event==i]))
}
```

# Stan - Generated quantities

```{r}
gen_time <- seq(1, max(as.numeric(modelDatUse$Time)))
gen_time_N <- length(gen_time)
```


# Initial estimates

```{r}
ini = function(){ #initial values for each parameter
  list(lambda_mu = abs(rnorm(1,10000,0.2)),
       d_mu = abs(rnorm(1,0.03,0.01)),
       k_mu = abs(rnorm(1,2.6e-8,1e-8)),
       delta_mu = abs(rnorm(1,0.3,0.2)),
       rho_mu = abs(rnorm(1,400,0.2)),
       c_mu = abs(rnorm(1,5,0.2)),
       k_phi = rep(abs(rnorm(1,2.6e-8,1e-8)), numIDs),
       #sigma = rep(abs(rnorm(1,0.5,0.2)), numIDs),
       #eta_lambda = rep(0.01, numIDs),
       #eta_d = rep(0.01, numIDs),
       #eta_k = rep(0.01, numIDs),
       #eta_delta = rep(0.01, numIDs),
       #eta_rho = rep(0.01, numIDs),
       #eta_c = rep(0.01, numIDs),
        sigma = rep(abs(rnorm(1,0.5,0.2)), 1)
       )
}

```


# Run model
```{r}
# Problem is ragged arrays for time and obs. Code group sizes
numIterations <- 400
modelName <- '3ODEMWE/ModelFitV2s12'
chains <- 2

fit <- stan(
            file = file.path(wd, paste(modelName, '.stan', sep = '')), 
            data=list(
                      nSubjects = numIDs, 
                      nObs = totNumObs, 
                      nODEs = numODEs, 
                      time = as.numeric(modelDatUse$Time),
                      obs = modelDatUse$VL, 
                      s = array(sizes),
                      t0 = t0, 
                      Y0s = array(Y0s),
                      gen_time = gen_time,
                      adapt_delta = 0.8,
                      max_treedepth = 10, #default = 10
                      gen_time_N = gen_time_N
                      #stepsize = 0.01, #when adaption is off
                      #engaged = 0 #0 = off
                      ),
            iter = numIterations,
            #init = ini,
            chains = chains,
            verbose = F,
            refresh = 1
            )

#print(fit, pars = c('thetaHat', 'sigma', 'U0', 'I0', 'lp__', 'eta'), probs = c(0.1, 0.5, 0.9), digits = 3)
#print(fit)
```

```{r}
print(fit)
```


# Generated Predictions
```{r}
n <- 3 #number of ODEs

generated_quantities <- extract(fit)

pop_u <- generated_quantities$'pop_pred'[,,1]
pop_i <- generated_quantities$'pop_pred'[,,2]
pop_v <- generated_quantities$'pop_pred'[,,3]

ind_u <- generated_quantities$'ind_pred'[,,1]
ind_i <- generated_quantities$'ind_pred'[,,2]
ind_v <- generated_quantities$'ind_pred'[,,3]

U0_hat <- generated_quantities$'Y0_hat'[,1,]
I0_hat <- generated_quantities$'Y0_hat'[,2,]
V0_hat <- generated_quantities$'Y0_hat'[,3,]


generated_quantities_dat <- modelDatUse

generated_quantities_dat$chain <- NA

generated_quantities_dat$pop_U <- NA
generated_quantities_dat$pop_I <- NA
generated_quantities_dat$pop_V <- NA

generated_quantities_dat$ind_U <- NA
generated_quantities_dat$ind_I <- NA
generated_quantities_dat$ind_V <- NA

generated_quantities_dat$U0_hat <- NA
generated_quantities_dat$I0_hat <- NA
generated_quantities_dat$V0_hat <- NA


generated_quantities_dat <- generated_quantities_dat[rep(seq_len(nrow(generated_quantities_dat)), dim(pop_v)[1]),] #create repeating dataframe to contain structured sample computations

#pop values
pop_vs <- vector()
pop_is <- vector()
pop_us <- vector()

ind_vs <- vector()
ind_is <- vector()
ind_us <- vector()

#U0_hats <- vector()
#I0_hats <- vector()
#V0_hats <- vector()

for (i in 1:dim(pop_v)[1]){ #loop through chains
 pop_vs <- append(pop_vs, pop_v[i,])
 pop_is <- append(pop_is, pop_i[i,])
 pop_us <- append(pop_us, pop_u[i,])
 
 ind_vs <- append(ind_vs, ind_v[i,])
 ind_is <- append(ind_is, ind_i[i,])
 ind_us <- append(ind_us, ind_u[i,])
 
 #U0_hats <- append(U0_hats, rep(U0_hat[i,], times = sizes))
 #I0_hats <- append(I0_hats, rep(I0_hat[i,], times = sizes))
 #V0_hats <- append(V0_hats, rep(V0_hat[i,], times = sizes))
}

generated_quantities_dat$pop_V <- pop_vs
generated_quantities_dat$pop_I <- pop_is
generated_quantities_dat$pop_U <- pop_us

generated_quantities_dat$ind_V <- ind_vs
generated_quantities_dat$ind_I <- ind_is
generated_quantities_dat$ind_U <- ind_us

#generated_quantities_dat$U0_hat <- U0_hats
#generated_quantities_dat$I0_hat <- I0_hats
#generated_quantities_dat$V0_hat <- V0_hats

generated_quantities_dat$chain <- rep(1:dim(pop_v)[1], each = dim(pop_v)[2])


```

```{r}
temp <- generated_quantities_dat[generated_quantities_dat$chain==max(generated_quantities_dat$chain),]

#plot(temp$V, temp$pop_V)

#qplot(temp$V, temp$pop_V) + geom_abline(intercept = 0, slope = 1)
qplot(temp$VL, temp$ind_V, colour = as.factor(temp$ID), group = as.factor(temp$ID)) + geom_abline(intercept = 0, slope = 1)+scale_y_log10(labels = comma)+scale_x_log10(labels = comma)
#qplot(temp$V, temp$pop_V, colour = as.factor(temp$ID), group = as.factor(temp$ID)) + geom_abline(intercept = 0, slope = 1)
#qplot(generated_quantities_dat$V, generated_quantities_dat$ind_V) + geom_abline(intercept = 0, slope = 1)
```

```{r}
#plot(pop_vl)
```


```{r}
#pairs(fit, pars = c('theta', 'sigma','U0', 'I0'))
#pairs(fit, pars = c('theta', 'sigma', 'U0'))
#pairs(fit, pars = c('theta', 'sigma'))
#pairs(fit, pars = c('theta'))
#pairs(fit, pars = c('lambda', 'delta', 'rho'))
#pairs(fit, pars = c('d', 'startingI'))
```

```{r}
#traceplot(fit, pars = c('theta', 'sigma', 'U0', 'I0'))
#traceplot(fit, pars = c('theta', 'sigma', 'U0'))
#traceplot(fit, pars = c('theta', 'sigma'))
#traceplot(fit, pars = c('theta', 'eta_beta'), inc_warmup = T)
traceplot(fit, inc_warmup = T, pars = c('lambda_mu', 'lambda_phi'))
```

```{r}
ggplot(data = generated_quantities_dat[generated_quantities_dat$chain == max(generated_quantities_dat$chain),])+
  geom_point(aes(x = Time, y = pop_V, group = ID, colour = as.factor(ID)))+
  geom_line(aes(x = Time, y = pop_V, group = ID, colour = as.factor(ID)))+
  scale_y_log10(labels = comma)
```


```{r}
ggplot(data = generated_quantities_dat[generated_quantities_dat$chain == max(generated_quantities_dat$chain),])+
  geom_point(aes(x = Time, y = ind_V, group = ID, colour = as.factor(ID)))+
  geom_line(aes(x = Time, y = ind_V, group = ID, colour = as.factor(ID)))+
  scale_y_log10(labels = comma)
```


```{r}
ggplot(data = generated_quantities_dat[generated_quantities_dat$chain == max(generated_quantities_dat$chain),])+
  geom_point(aes(x = ind_V, y = pop_V, group = Event, colour = as.factor(Event)))+
  geom_line(aes(x = ind_V, y = pop_V, group = Event, colour = as.factor(Event)))+
  scale_y_log10(labels = comma)+
  scale_x_log10(labels = comma)
```

```{r}
ggplot(data = generated_quantities_dat[generated_quantities_dat$chain == max(generated_quantities_dat$chain),])+
  geom_point(aes(x = ind_V, y = pop_V, group = Event, colour = as.factor(Event)))+
  #geom_line(aes(x = ind_V, y = pop_V, group = Event, colour = as.factor(Event)))+
  scale_y_log10(labels = comma)+
  scale_x_log10(labels = comma)
```


```{r}
ggplot(data = generated_quantities_dat[generated_quantities_dat$Event %in% c(134, 256, 94, 200, 339, 407, 461) & generated_quantities_dat$chain > 1000,])+
  geom_line(aes(x = Time, y = ind_V, group = interaction(chain, Event), colour = as.factor(Event)), alpha = .01)+
  geom_point(aes(x = Time, y = ind_V, group = interaction(chain, Event), colour = as.factor(Event)), alpha = .1)+
  scale_y_log10(labels = comma, breaks = c(10, 100, 1000, 10000, 100000, 1000000))+
  #scale_x_log10(labels = comma, breaks = c(10, 100, 1000, 10000, 100000, 1000000))+
  theme_classic()+
  #geom_abline(intercept = 0, slope = 1)+
  labs(x = 'Time Post Transplant', y = 'Individual Predictions')+
  theme(legend.position = 'none')
```


```{r}
ggplot(data = generated_quantities_dat[generated_quantities_dat$chain == max(generated_quantities_dat$chain),])+
  geom_point(aes(x = Time, y = ind_V, group = Event, colour = as.factor(Event)), shape = 2)+
  geom_line(aes(x = Time, y = ind_V, group = Event, colour = as.factor(Event)), linetype = 2)+
  geom_point(aes(x = Time, y = VL, group = Event, colour = as.factor(Event)))+
  
  geom_point(aes(x = -10, y = V0_hat, group = Event, colour = as.factor(Event)))+
  
  #geom_point(aes(y = generated_quantities_dat$U0_hat[generated_quantities_dat$chain == max(generated_quantities_dat$chain) & generated_quantities_dat$Time == 0], x = -1, group = generated_quantities_dat$Event[generated_quantities_dat$chain == max(generated_quantities_dat$chain) & generated_quantities_dat$Time == 0], colour = as.factor(generated_quantities_dat$Event[generated_quantities_dat$chain == max(generated_quantities_dat$chain) & generated_quantities_dat$Time == 0])), shape = 3)+
  
  geom_line(aes(x = Time, y = VL, group = Event, colour = as.factor(Event)))+
  scale_y_log10(labels = comma)+
  #scale_x_log10(labels = comma)+
  NULL
```

