---
title: "ODE Simulations"
author: "Ben Margetts"
date: "11/04/2017"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

# Purpose: Simulate ODE Version of Model. Make Sure it Looks OK.

# Required Packages

Load in the following packages:

```{r results='hide', message=FALSE, warning=FALSE}
library(Rcpp)
library(data.table)
library(zoo)
library(plyr)
library(scales)
library(ggplot2)
library(ggthemes)
library(dplyr)
library(plotly)
library(pracma)
library(Hmisc)
#library(rpart)
#library(survminer)
library(knitr)
library(rstan)
#library(shinystan)
library(deSolve)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Setup notebook

```{r setup}
opts_chunk$set(message=T, warning=T, tidy.opts=list(width.cutoff=60),tidy=F)
wd = "~/Google/PhD/Year_1/Data/BMT_Data/BMT_Drug_Dec2015" #Working directory variable
```

```{r}
# Set up the true parameter values
a <- 0.8
b <- 2
sigma <- .2
     
# Simulate some data
x <- (1:1000)/100
N <- length(x)
obs <- a*exp(b*x)
```


#Ordinary Differential Equation 1 Vset - 1 Comp.

```{r}
#Without Noise
t0 <- 0
Y0 <- 100
numODEs <- 1
#Create the ODE
Exp <- function(t, state, parameters) {
with(as.list(c(state, parameters)), { 
  dY<- delta*Y * (1-(Y/yset)) - c*Y
  list(dY)
      })
}
parameters <- c(delta = 3, c=1, yset=50000
                )
state <- c(Y=Y0)
times <- (1:1000)/100
```

We can then run the ODE that we set up above using the following code.

```{r}
#Run the ODE
out <- ode(y = state, times = times, func = Exp, parms = parameters)
```

If we plot the results of the analytical solution of our function against the results from our ODE, we can see that they agree.

```{r}
plot(out[,1],out[,2])
```


#Ordinary Differential Equation 2 - 3 Comp.


```{r}
t0 <- 0
Y0 <- 100
numODEs <- 3
#Create the ODE
Exp <- function(t, state, parameters) {
with(as.list(c(state, parameters)), {
  
  dU <- (-beta*U*V) + (lambda)
  dI <- (beta*U*V) - (delta*I)
  dV <- (rho*delta*I) - (c*V)
  list(c(dU, dI, dV))
      
  })
}
#parameters <- c(beta=1, lambda=4000, delta=0.1, rho=1, c=1 
#                )
#state <- c(U=7441321, I=0, V=100)
state <- c(U=100, I=0, V=100)
#times <- (1:max(testDat$DATE))/1000
times <- 1:100
```

We can then run the ODE that we set up above using the following code.

```{r}
#Run the ODE
numPatients <- 3
out <- vector()
# remove this line
#parameters <- c(beta=runif(1, min = 0.0001, max = 0.1), lambda=runif(1), delta=runif(1), rho=runif(1), c=runif(1))
parameters <- c(beta = 1, lambda = 1, delta = 1, rho = 1, c = 1)
for (i in 1:numPatients){
  parameters <- c(beta = 1, lambda = 1, delta = 0.5, rho = 5, c = 1)
  #parameters <- c(beta=runif(1, min = 0.0001, max = 0.1), lambda=runif(1), delta=runif(1), rho=runif(1), c=runif(1))
  #times <- 1:runif(1, min = 10, max = 365)
  #parameters <- parameters + 0.05
  parameters <- parameters + rnorm(1, 0, 0.1)
  
  V0 <- runif(1, min = 1000, max = 10000)
  
  state <- c(U=100, I=0, V=V0)
  times <- 1:100
  
  if (length(out) != 0){
    temp <- ode(y = state, times = times, func = Exp, parms = parameters)
    temp <- as.data.frame(temp)
    temp$ID <- i
    temp$beta <- parameters[1]
    temp$lambda <- parameters[2]
    temp$delta <- parameters[3]
    temp$rho <- parameters[4]
    temp$c <- parameters[5]
    out <- rbind(out,temp)
  }
  
    if (length(out) == 0){
    out <- ode(y = state, times = times, func = Exp, parms = parameters)
    out <- as.data.frame(out)
    out$ID <- i
    out$beta <- parameters[1]
    out$lambda <- parameters[2]
    out$delta <- parameters[3]
    out$rho <- parameters[4]
    out$c <- parameters[5]

  }

}
```

If we plot the results of the analytical solution of our function against the results from our ODE, we can see that they agree.

```{r}
plot(out[,1],out[,3])
```

```{r}
plot(out[,1], out[,4])
```

```{r}
setwd('~/Google/PhD/Year_1/Modelling/Stan/CMV Stan Feb 18')

write.csv(out, 'SimulatedData2.csv')
```

